

\preClass{Second order, constant coefficient equations.}

  \begin{problem}

  \item A spring mass system is constructed. The mass is 0.1 kg. The
    spring is stretched 0.04 m when the mass is hung vertically from
    the spring. The force due to air friction is estimated to be 0.36
    N when the object is moving at 0.2 m/sec. The system is placed in
    a horizontal orientation, and the object is pulled so that the
    spring extends 0.03 m and released from rest. (Assume that $g$ is
    10 m/sec\textsuperscript{2}.)

    \begin{subproblem}
      \item Sketch the free body diagram.
        \vspace{4em}

      \item Determine the governing equation for the system. (This
        includes the initial conditions.)
        \vfill


      \item Determine the \textit{general} solution to the differential
        equation. (Do not consider the initial conditions yet.)
        \vfill

        \clearpage

      \item Determine the specific solution to match the initial conditions.

        \vfill

    \end{subproblem}



  \end{problem}


\actTitle{Second order, linear differential equations}
\begin{problem}
\item A spring mass system will be constructed. The mass is $m$
  kg. The spring is stretched 0.04 m when a test mass of $0.1$ kg 
  is hung vertically from the spring. The air friction coefficient is $b$.

    \begin{subproblem}
      \item Determine the governing equation for the system. 
        \vfill

      \item Determine the characteristic equation for the differential
        equation. (Do not consider the initial conditions yet.)
        \vfill

        \clearpage

      \item Assume that the solution to the characteristic equation
        will be a complex number so that the system will
        oscillate. Break the solutions into two pieces, the real and
        the imaginary parts.

        \vfill

      \item Find and solve the system of equations that result for $m$
        and $b$ if you wish the amplitude of the system to decay like
        $e^{-t/10}$ and the system to oscillate with a period of 3.1
        seconds.

        \vfill

    \end{subproblem}


\end{problem}



\actTitle{Method of Undetermined Coefficients}

  \begin{problem}
  \item In the pre-class work you had to determine the position of the
    mass in a spring mass system. We now examine that result.
    \begin{subproblem}
    \item Write out the solution to the differential equation found in
      the pre-class work.
      \vspace{2em}
    \item The kinetic energy of a mass is $\frac{1}{2} m v^2$, where
      $m$ is the mass of the object and $v$ is the speed. Find the
      kinetic energy of the mass at any time.
      \vfill

    \clearpage

  \item The potential energy stored in a spring is $-\frac{1}{2} k h |h|$
    where $k$ is the spring constant and $h$ is the displacement of
    the spring. (If $h$ is negative the spring is under compression.)
    Determine the potential energy of the spring at any time.
    \vfill

    \clearpage
  \item The total energy of the spring is the potential energy plus
    the kinetic energy. Determine the total energy in the system at
    any time. 

    \vfill

  \item Sketch a plot for the total energy as a function of time.
    \vspace*{8em}


  \end{subproblem}



\end{problem}

