


\preClass{Second order, constant coefficient equations.}

\begin{problem}
\item Suppose that a function is a polynomial of degree 2,
  \begin{eqnarray*}
    p(t) & = & 4t^2 + 3t - 1
  \end{eqnarray*}

  \begin{subproblem}
  \item Determine the first derivative of $p(t)$. What degree polynomial is
    the derivative?

    \vspace{4cm}

  \item Determine the second derivative of $p(t)$. What degree
    polynomial is the second derivative?  

    \vspace{4cm}

  \item Determine the value of $p''(t)+3p'(t)+2p(t)$. What is the
    degree of this new polynomial? 

    % 8 t^2 + 30t + 15

    \vfill

  \end{subproblem}

\end{problem}


\actTitle{Second Order Equations}
  \begin{problem}

    \item Determine the solution to the differential equation
      \begin{eqnarray*}
        x''(t) + 3x'(t) + 2x(t) & = & 8 t^2 + 30t + 15.
      \end{eqnarray*}

      \begin{subproblem}
        \item First, determine the homogeneous solution to the
          equation.

          \vfill

        \item Now assume a general form for the particular
          solution. Here we are looking for a function, $p(t)$, so
          that 
          \begin{eqnarray}
            p''(t) + 3p'(t) + 2p(t) & = & 8 t^2 + 30t + 15.
            \label{eqn:quadraticParticular}
          \end{eqnarray}
          If $p(t)$ is a quadratic polynomial then we would expect the
          left hand side to be equal to a quadratic polynomial.

          Express $p(t)$ as a general quadratic polynomial.

          \vspace*{4em}

        \item Find the first and second derivatives of the general
          form of $p(t)$.

          \vfill

          \clearpage

        \item Substitute the previous results into equation
          \ref{eqn:quadraticParticular}.

          \vfill

        \item Simplify the expression by collecting all the terms to
          one side and collect the constant, linear, and quadratic
          terms together.

          \vfill

        \item Set each coefficient to zero and determine the three
          equations that must be true.

          \vfill

          \clearpage

        \item Determine the particular solution.
          \vfill

        \item Using the homogeneous and particular solutions determine
          the solution to the differential equation.

          \vfill

      \end{subproblem}

      \clearpage

    \item Determine the solution to the differential equation
      \begin{eqnarray*}
        x''(t) + 3x'(t) + 2x(t) & = & -10 t^3 - 45 t^2 - 32 t + 5.
      \end{eqnarray*}

      \vfill
      \clearpage


    \item Determine the solution to the differential equation
      \begin{eqnarray*}
        x''(t) + 3x'(t) + 2x(t) & = & -10 t^3.
      \end{eqnarray*}

      \vfill
      \clearpage

    \item Determine the solution to the differential equation
      \begin{eqnarray*}
        x''(t) + 3x'(t) + 2x(t) & = & e^{4t}.
      \end{eqnarray*}

      \vfill


      \clearpage


\end{problem}

\actTitle{Second Order Equations}
\begin{problem}

\item Now we solve the problem a different way. Recall the
  differential equation,
  \begin{eqnarray}
    x''(t) + 3x'(t) + 2x(t) & = & 8 t^2 + 30t + 15.
    \label{eqn:secondOrderExample}
  \end{eqnarray}

  \begin{subproblem}
    \item Determine the homogeneous solutions:
      \vspace*{4em}

    \item Using your homogeneous solutions, $y_1$ and $y_2$, express
      the solution as
    \begin{eqnarray*}
      y(t) & = & y_1(t) u_1(t) + y_2(t) u_2(t).
    \end{eqnarray*}
    The functions $u_1$ and $u_2$ are unknown, and our goal is to find
    new equations to help us determine a formula for these two
    functions.  Substitute your values for $y_1$ and $y_2$ as the
    homogeneous solutions.

    \vspace{4em}

  \item Determine the first derivative of $y(t)$ above using the
    product rule.

    \vfill

  \item Assume that $y_1(t) u_1'(t) + y_2(t) u_2'(t)$ is equal to
    zero. Write out the new equation for $y(t)$ using this assumption.

    \vfill

  \item Determine the second derivative of $y(t)$ above using the
    product rule and the previous result.

    \vfill

    \clearpage

  \item Substitute your results for the function $y(t)$ and its
    derivatives into the differential equation, equation
    (\ref{eqn:secondOrderExample}).

    \vfill

  \item Write out the two equations that you have. (The first equation
    comes from assuming $y_1(t) u_1'(t) + y_2(t) u_2'(t)=0$.

    \vfill

  \item Solve the equations for $u_1'(t)$ and $u_2'(t)$.

    \vfill

    \clearpage

  \item Solve the integrals to determine $u_1$ and $u_2$.

    \vfill

  \item Determine the solution to the differential equation.

    \vfill


  \end{subproblem}

  \clearpage

\item Repeat the previous steps for the equation
  \begin{eqnarray*}
    x''(t) + 3x'(t) + 2x(t) & = & e^{4t}.
  \end{eqnarray*}


\end{problem}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
