  \actTitle{Second Order Equations}
  \begin{problem}
  \item The following questions refer to the following differential
    equation:
    \begin{eqnarray*}
      y'' - y - 12 = 0.
    \end{eqnarray*}

    \begin{subproblem}
      \item Show that $y=e^{4t}$ is a solution.
        \vspace{4cm}
      \item Show that $y=e^{-3t}$ is a solution.
        \vspace{4cm}
      \item Show that $y=C_1 e^{4t} + C_2 e^{-3t}$ is a solution for
        any constants $C_1$ and $C_2$. Will the solution grow or
        decay?

        \vfill

    \end{subproblem}
 
    \clearpage
  \item Find the solutions to the following differential equation:
    \begin{eqnarray*}
      y'' + 3y' - 10y = 0.
    \end{eqnarray*}
    Determine if the solution will exhibit growth or decay. Can you
    determine the answer to the question of stability by only looking
    at the roots of the characteristic equation? If so how?

    \vfill
  \end{problem}


  \actTitle{Second Order Equations}
  \begin{problem}

  \item The questions refer to the following differential equation:
    \begin{eqnarray*}
      y'' + 4y & = & 0.
    \end{eqnarray*}

    \begin{subproblem}
      \item Show that $y=C_1 e^{i2t} + C_2 e^{-i2t}$ is a solution to
        the differential equation.
        \vfill

      \item Use Euler's formula to write the solution, $y=C_1 e^{i2t}
        + C_2 e^{-i2t}$, in terms of sines and cosines.
        \vfill

        \clearpage

      \item Simplify the terms so that the sine and cosine terms are
        only expressed by either $\sin(2t)$ or $\cos(2t)$.

        \vfill

      \item Collect all the sine terms and the cosine terms so that
        the solution is in the form
        \begin{eqnarray*}
          y & = & \#_1 \cos(2t) + \#_2 \sin(2t).
        \end{eqnarray*}
        \vfill

      \item Show that $y=k_1 \cos(2t) + k_2 \sin(2t)$ is a solution to
        the original differential equation.
        \vfill
        

    \end{subproblem}


\end{problem}
