  \actTitle{Second order, linear differential equations}
  \begin{problem}

  \item A spring mass system is constructed. The mass is 0.1 kg. The
    spring is stretched 0.04 m when the mass is hung vertically from
    the spring. The force due to air friction is estimated to be 0.03
    N when the object is moving at 0.2 m/sec. The system is placed in
    a horizontal orientation, and the object is pulled so that the
    spring extends 0.03 m and released from rest.

    \begin{subproblem}
      \item Determine the governing equation for the system. (This
        includes the initial conditions.)
        \vfill

      \item Determine the \textit{general} solution to the differential
        equation. (Do not consider the initial conditions yet.)
        \vfill

        \clearpage

      \item Determine the specific solution to match the initial conditions.

        \vfill

      \item Determine a different value for the spring constant so
        that the system will oscillate with a period of 1.2 seconds.
        \vfill

    \end{subproblem}



  \end{problem}



\actTitle{Method of Undetermined Coefficients}

  \begin{problem}
  \item Find the solution to the following differential equation:
    \begin{eqnarray*}
      y'' - 2y' - 8y & = & 3t + 4.
    \end{eqnarray*}
    \begin{subproblem}
    \item Write out the homogeneous equation.
      \vspace{2em}
    \item Find the solution to the homogeneous equation.
      \vfill

    \clearpage

  \item Write out the general form of the particular solution:
    \vspace{2em}

  \item Substitute your particular solution into the non-homogeneous equation.
    \vfill

  \item Solve for the constants.
    \vfill

  \end{subproblem}

  \end{problem}

